Static and Dynamic Bethe-Salpeter Equations in the $T$-Matrix Approximation

### 2022-02-09 Manuscript/Tmatrix.tex
saving work b4 transfer to overleaf

In terms of computational cost, the overall scaling of BSE@$GT$ is equivalent to BSE@$GW$ as they both correspond to seeking the lowest eigenvalues of a matrix of size $(2OV \times 2OV)$. 
Searching iteratively for the lowest eigenstates can be routinely performed via Davidson's algorithm with a $\order*{K^4}$ computational cost. \cite{Davidson_1975}
The cost of the dynamical correction, which is thoroughly discussed in Ref.~\onlinecite{Loos_2020h}, is more expensive but is again equivalent in both formalisms.
The computational cost associated with the computation of the $T$-matrix and the screening $W$ both scale as $\order*{K^6}$ as one must obtain all the eigenvalues and eigenvectors of the pp-RPA and ph-RPA problem, respectively.
However, the prefactor of the pp-RPA calculation is significantly larger than its ph-RPA counterpart due to the larger size of the pp-RPA matrices and its non-Hermitian nature. \cite{Peng_2013,Scuseria_2013,Yang_2013,Yang_2013b,Yang_2014a,Zhang_2015,Zhang_2016}
Moreover, within the $T$-matrix formalism, one must compute both the singlet and triplet contributions of the $T$-matrix, while for singlet states, only the singlet contribution of $W$ contributes to the self-energy.
Although similar approaches remain to be developed for the $T$-matrix formalism, contour deformation and density fitting techniques can be efficiently implemented in the case of $GW$ to reduce the scaling to $\order*{K^3}$. \cite{Duchemin_2019,Duchemin_2020,Duchemin_2021}

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Excited states of the hydrogen molecule}
\label{sec:h2}
%%%%%%%%%%%%%%%%%%%%%%
As a first didactical example, we consider the excited states of the hydrogen molecule \ce{H2} and the variation of the excitation energies upon dissociation.
The excitation energies for some low-lying singlet and triplet excited states are represented in Fig.~\ref{fig:H2} as a function of the internuclear distance $R_{\ce{H-H}}$ at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels with the cc-pVTZ basis.

@@

%%% FIGURE 2 %%%
\begin{figure}
	\includegraphics[width=\linewidth]{H2_gap}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed at the BSE@$G_0W_0$ (blue) and BSE@$G_0T_0$ (red) levels with the cc-pVTZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:H2}
\end{figure}
%%% %%% %%% %%%

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Excited states of beryllium hydride}
\label{sec:beh2}
%%%%%%%%%%%%%%%%%%%%%%

%%% FIGURE 3 %%%
\begin{figure*}
	\includegraphics[width=\textwidth]{BeH2}
	\caption{Singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{BeH2} as a function of the distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels.
	The dynamically-corrected BSE excitation energies are represented as thin lines for dBSE@$G_0W_0$ (blue), and dBSE@$G_0T_0$ (red) with the cc-pVDZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:H2}
\end{figure*}
%%% %%% %%% %%%

%%% FIGURE 4 %%%
\begin{figure}
	\includegraphics[width=\linewidth]{BeH2_gap}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{BeH2} as a function of the internuclear distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the BSE@$G_0W_0$ (blue) and BSE@$G_0T_0$ (red) levels with the cc-pVDZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:H2}
\end{figure}
%%% %%% %%% %%%

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Diatomics}
\label{sec:diatomics}
%%%%%%%%%%%%%%%%%%%%%%

%%% TABLE I %%%
\begin{table*}
	\caption{Singlet and triplet excitation energies (in \si{\eV}) of various molecular systems at equilibrium geometry computed at various levels of theory with the cc-pVDZ basis set.
	The corresponding experimental equilibrium geometries have been extracted from the computational chemistry comparison and benchmark database (CCCBDB) at \url{https://cccbdb.nist.gov}.}
	\label{fig:mol}
	\begin{ruledtabular}
		\begin{tabular}{lcccccccc}
					&			&	\mc{7}{c}{Methods}																\\
			\cline{3-9}
			Mol.	&	State	&	CIS		&	CIS(D)	&	TDHF	&	BSE@$GW$	&	dBSE@$GW$	&	BSE@$GT$	&	dBSE@$GT$	\\
			\hline
			\ce{H2}	&			&	\\
			\ce{LiH}	&			&	\\
			\ce{LiF}	&			&	\\
			\ce{BF}	&			&	\\
			\ce{CO}	&			&	\\
			\ce{N2}	&			&	\\
			\ce{F2}	&			&	\\
			\ce{FH}	&			&	\\
		\end{tabular}
	\end{ruledtabular}
\end{table*}
%%% %%% %%% %%%

>>>>

In terms of computational cost, the overall scaling of BSE@$GT$ is equivalent to BSE@$GW$ as they both correspond to seeking the lowest eigenvalues of a matrix of size $(2OV \times 2OV)$. 
Searching iteratively for the lowest eigenstates can be routinely performed via Davidson's algorithm with a $\order*{K^4}$ computational cost. \cite{Davidson_1975}
The cost of the dynamical correction, which is thoroughly discussed in Ref.~\onlinecite{Loos_2020h}, is more expensive but is again equivalent in both formalisms.
The computational cost associated with the computation of the $T$-matrix and the screening $W$ both scale as $\order*{K^6}$ as one must obtain, respectively, all the eigenvalues and eigenvectors of the pp-RPA and ph-RPA problems.
However, the prefactor of the pp-RPA calculation is significantly larger than its ph-RPA counterpart due to the larger size of the pp-RPA matrices and its non-Hermitian nature. \cite{Peng_2013,Scuseria_2013,Yang_2013,Yang_2013b,Yang_2014a,Zhang_2015,Zhang_2016}
Moreover, within the $T$-matrix formalism, one must compute both the singlet and triplet contributions of the $T$-matrix, while for singlet states, only the singlet contribution of $W$ contributes to the self-energy.
Although similar approaches remain to be developed for the $T$-matrix formalism, contour deformation and density fitting techniques can be efficiently implemented in the case of $GW$ to reduce the scaling to $\order*{K^3}$. \cite{Duchemin_2019,Duchemin_2020,Duchemin_2021}

@@

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Excited states of the hydrogen molecule}
\label{sec:h2}
%%%%%%%%%%%%%%%%%%%%%%
As a first didactical example, we consider the lowest singlet and triplet excited states of the hydrogen molecule \ce{H2} and the variation of their respective excitation energies upon dissociation.
The excitation energies associated with these low-lying singlet and triplet excited states are represented in Fig.~\ref{fig:H2} as a function of the internuclear distance $R_{\ce{H-H}}$ at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels with the cc-pVTZ basis.
The variation of the HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap computed at the $G_0W_0$ and $G_0T_0$ levels is depicted in Fig.~\ref{fig:H2_gap}

@@

%%% FIGURE 2 %%%
\begin{figure}
	\includegraphics[width=\linewidth]{H2_gap}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{H2} as a function of the internuclear distance $R_{\ce{H-H}}$ (in \si{\angstrom}) computed at the $G_0W_0$ (blue) and $G_0T_0$ (red) levels with the cc-pVTZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:H2_gap}
\end{figure}
%%% %%% %%% %%%

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Excited states of beryllium hydride}
\label{sec:beh2}
%%%%%%%%%%%%%%%%%%%%%%
As a second example, we consider the symmetric dissociation of the linear molecule beryllium hydride \ce{BeH2}.
The variation in excitation energies as a function of the distance $R_{\ce{Be-H}}$ is shown in Fig.~\ref{fig:BeH2}.
The variation of the HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap computed at the $G_0W_0$ and $G_0T_0$ levels is depicted in Fig.~\ref{fig:BeH2_gap}

%%% FIGURE 3 %%%
\begin{figure*}
	\includegraphics[width=\textwidth]{BeH2}
	\caption{Singlet (left) and triplet (right) excitation energies (in \si{\eV}) of \ce{BeH2} as a function of the distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the FCI (black), BSE@$G_0W_0$ (blue), and BSE@$G_0T_0$ (red) levels.
	The dynamically-corrected BSE excitation energies are represented as thin lines for dBSE@$G_0W_0$ (blue), and dBSE@$G_0T_0$ (red) with the cc-pVDZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:BeH2}
\end{figure*}
%%% %%% %%% %%%

%%% FIGURE 4 %%%
\begin{figure}
	\includegraphics[width=\linewidth]{BeH2_gap}
	\caption{HOMO and LUMO quasiparticle energies as well as HOMO-LUMO gap (in \si{\eV}) of \ce{BeH2} as a function of the internuclear distance $R_{\ce{Be-H}}$ (in \si{\angstrom}) computed at the BSE@$G_0W_0$ (blue) and BSE@$G_0T_0$ (red) levels with the cc-pVDZ basis.
	Raw data are reported in {\SupInf}.}
	\label{fig:BeH2_gap}
\end{figure}
%%% %%% %%% %%%

%%%%%%%%%%%%%%%%%%%%%%
\subsection{Molecules at equilibrium}
\label{sec:mol}
%%%%%%%%%%%%%%%%%%%%%%

Table \ref{tab:mol} reports singlet and triplet vertical excitation energies for various molecular systems at equilibrium geometry.
In addition to the BSE-based models studied in the present manuscript, we have selected well-known methods \titou{(CIS, CIS(D), TDHF and FCI)}.

%%% TABLE I %%%
\begin{table*}
	\caption{Singlet and triplet excitation energies (in \si{\eV}) of various molecular systems at equilibrium geometry computed at various levels of theory with the cc-pVDZ basis set.
	The corresponding experimental equilibrium geometries have been extracted from the computational chemistry comparison and benchmark database (CCCBDB) at \url{https://cccbdb.nist.gov}.}
	\label{tab:mol}
	\begin{ruledtabular}
		\begin{tabular}{lcccccccc}
					&			&	\mc{7}{c}{Methods}																\\
			\cline{3-9}
			Mol.	&	State	&	CIS		&	CIS(D)	&	TDHF	&	BSE@$GW$	&	dBSE@$GW$	&	BSE@$GT$	&	dBSE@$GT$	\\
			\hline
			\ce{H2}	&			&	\\
			\ce{LiH}	&			&	\\
			\ce{LiF}	&			&	\\
			\ce{BF}	&			&	\\
			\ce{CO}	&			&	\\
			\ce{N2}	&			&	\\
			\ce{F2}	&			&	\\
			\ce{FH}	&			&	\\
		\end{tabular}
	\end{ruledtabular}
\end{table*}
%%% %%% %%% %%%
